---
title: "Comparison of tracts after registration"
output: html_notebook
---

## Intro

The idea is to try and compare the same FlyCircuit neurons after registration
using CMTK (Cambridge) or FlyCircuit's default pipeline (FCv1.2 online).

It is my impression going into this analysis that the mean and absolute
dispersion after the Amira based registration in v1.2 is greater than for our
standard CMTK registration. I would put this down to the use of single 
template brains that were acquired some time ago rather than an average brain,
which has higher S/N and is more medial in its overall shape. See also 
Bogovic and Saalfeld 


```{r}
library(flycircuit)
library(natverse)
```

```{r}
load("dl2d.rda")
```

```{r}
clear3d()
plot3d(dl2d.both, col=orig, colpal = c('cam'='cyan', fc='red'))
```


```{r}
plot(dl2d.fc)
```

Let's cut around 400

```{r}
dl2d.both=c(dl2d.cam, dl2d.fc)
dl2d.both[,'orig']=ifelse(names(dl2d.both) %in% names(dl2d.cam), 'cam', 'fc')

spines=nlapply(dl2d.both, spine)

tract_vector <- function(n, xval=400, thresh=10) {
  p=xyzmatrix(n)
  near=abs(p[,'X']-xval)<thresh
  pc=prcomp(p[near,], )
  c(colMeans(p[near,]), pc$rotation[,'PC1'])
}


tvsl=nlapply(spines, tract_vector, OmitFailures = T)
tvs=do.call(rbind, tvsl)
mean_cent=colMeans(tvs[,1:3])
mean_vec=colMeans(tvs[,4:6])


# nb *10 to make the vector longer (10 µm) and more visible
res2=rbind(mean_cent-mean_vec*10, mean_cent, mean_cent+mean_vec*10)

plot(spines, col='grey', WithNodes=F)
lines(res2[,1:2], lwd=3)

```

```{r}
plot3d(spines)
plc=plane_coefficients(mean_cent, mean_vec)
planes3d(plc[,1:3], d=plc[,'d'])

```

```{r}
clear3d()
make_centred_plane <- function(point, normal, scale = 1) {
  # find the two orthogonal vectors to this one
  uv = Morpho::tangentPlane(normal)
  uv = sapply(uv, "*", scale, simplify = F)
  qmesh3d(
  cbind(
  point + uv$z,
  point + uv$y,
  point - uv$z,
  point - uv$y
  ),
  homogeneous = F,
  indices = 1:4
  )
}
plane=make_centred_plane(mean_cent, mean_vec, scale=15)
plot3d(spines, col='grey')
shade3d(plane, col='red')
```

```{r}
intersections=t(sapply(dl2d.both, intersect_plane, plc, closestpoint=mean_cent))
intersections.cent=scale(intersections, center = T, scale = F)
d=sqrt(rowSums(intersections.cent^2))
mean(d)
```

```{r}
spheres3d(intersections, col='red', rad=.5)
shade3d(plane, col='black')
plot3d(spines, col='grey')
```

```{r}
uv = Morpho::tangentPlane(mean_vec)
# centre our points on mean
sp=scale(intersections, scale = F, center=T)
DotProduct=function(a,b) sum(a*b)
# find coordinates w.r.t. to our two basis vectors
xy=data.frame(u=apply(sp, 1, DotProduct, uv[[1]]),
           v=apply(sp, 1, DotProduct, uv[[2]]))
# maintain consist x-y aspect ratio 
plot(xy, pch=19, asp=1, main = "Position of Axon intersection in plane", 
     xlab="u /µm", ylab="v /µm")

```

```{r}
xy$orig=spines[,'orig']
plot(xy[1:2], pch=19, asp=1, main = "Position of Axon intersection in plane", 
     xlab="u /µm", ylab="v /µm", col=ifelse(xy$orig=='fc', 'red', '#a3c1ad'))

```


Let's look at dispersion by registration origin

```{r}
zz=by(xy[1:2], INDICES = xy$orig, scale, scale=FALSE)

rownorm <- function(x) apply(x, 1, function(r) sqrt(sum(r^2)))

devs=sapply(zz, rownorm, simplify = F)
sapply(devs, mean)
t.test(devs$cam, devs$fc)
```


Let's find a bunch of tract vectors near specific points.

```{r}
tract_vector_xyz <- function(n, xyz, thresh=10) {
  p=xyzmatrix(n)
  cent=scale(p, center = xyz, scale = F)
  d=sqrt(rowSums(cent^2))
  near=d<thresh
  pc=prcomp(p[near,, drop=FALSE], )
  c(colMeans(p[near,]), pc$rotation[,'PC1'])
}
tract_vectors <- function(x, xyz, ...) {
  tvsl=nlapply(spines, tract_vector_xyz, xyz=xyz, OmitFailures = T)
  tvs=do.call(rbind, tvsl)
  list(mean_cent=colMeans(tvs[,1:3]),
       mean_vec=colMeans(tvs[,4:6]))
}
```

Now to choose the specific points, let's take one of the spines and choose
evenly spaced points along it. First lets's choose the most medial spine.

```{r}
dl2d.both.dps=dotprops(spines, k=5, resample=1)
dl2d.both.aba=nblast_allbyall(dl2d.both.dps, normalisation = 'mean')
hist(colMeans(dl2d.both.aba))
best_neuron=names(which.max(colMeans(dl2d.both.aba)))
```


```{r}
spines[[best_neuron]]
best.30=resample(spines[[best_neuron]], stepsize = 30)
plot(best.30)
```


Now let's use the 3rd to n-1 points

```{r}
pointstouse <- xyzmatrix(best.30)[3:(nvertices(best.30)-2), ]
pointstouse
```

```{r}
planes=apply(pointstouse, 1, function(xyz) tract_vectors(spines, xyz=xyz, threshold=10))
head(planes, n = 3)
```


Now let's find the intersections on a plane for each of those locations

```{r}

project2plane <- function(data, mean_cent, mean_vec) {
  plc=plane_coefficients(mean_cent, mean_vec)
  intersections=t(sapply(data, intersect_plane, plc, closestpoint=mean_cent))
  uv = Morpho::tangentPlane(mean_vec)
  # centre our points on mean
  sp=scale(intersections, scale = F, center=T)
  DotProduct=function(a,b) sum(a*b)
# find coordinates w.r.t. to our two basis vectors
  xy=data.frame(u=apply(sp, 1, DotProduct, uv[[1]]),
           v=apply(sp, 1, DotProduct, uv[[2]]))
  xy
}

projected=sapply(planes, function(plc) project2plane(spines, plc$mean_cent,plc$mean_vec), simplify = F)
resdf=dplyr::bind_rows(projected, .id = 'pos')
resdf$name=rep(names(spines), length(projected))
resdf$orig=rep(spines[,'orig'], length(projected))
```

```{r}
library(ggplot2)
qplot(u,v, data=subset(resdf, pos%in%3:7), col=orig)+facet_wrap(~pos)+
  scale_color_manual(values = c('cam'='cyan', fc='red'))
```

Let's look at distance from centroid at each of those points

```{r}
library(dplyr)
resdf2 <- resdf %>%
  filter(pos%in%3:7) %>%
  group_by(pos, orig) %>%
  mutate(u=u-mean(u), v=v-mean(v)) %>%
  rowwise() %>%
  mutate(d=sqrt(sum(u^2+v^2)))

resdf3 <- resdf2 %>%
  group_by(pos, orig) %>%
  summarise(md=mean(d))
```

Now we can plot the mean dispersion at those different locations 

```{r}
qplot(factor(orig), md, data=resdf3, col=pos, ylim=c(0,NA), 
      ylab='Mean dispersion /µm', xlab='Registration')
```

```{r}
resdf3.wide=tidyr::pivot_wider(resdf3, names_from = 'orig', values_from='md')
resdf3.wide
```

```{r}
qplot(cam, fc, data=resdf3.wide, xlim=c(1.5,4.5), ylim=c(1.5,4.5),
      xlab='Mean dispersion, Cambridge /µm', ylab='Mean dispersion, FC 1.2 /µm')+
  geom_abline(slope=1, lty=2)+
  geom_smooth(method = 'lm', se = F)
```


```{r}
qplot(factor(orig), d, data=resdf2, col=orig) + 
  scale_colour_manual(values = c('cam'='cyan', fc='red'))+
  facet_wrap(~pos)+geom_boxplot()
```


```{r}
plot_centred_plane <- function(plc, scale=1, ...) {
  plane=make_centred_plane(plc$mean_cent, plc$mean_vec, scale=scale)
  shade3d(plane, ...)
}
```

```{r}
clear3d()
plot3d(spines[[best_neuron]], lwd=2, col='black')
# plot3d(spines, col=orig, colpal = c('cam'='#a3c1ad', fc='red'))
plot3d(spines, col=orig, colpal = c('cam'='cyan', fc='red'))
sapply(planes[-(1:2)], plot_centred_plane, scale=15, col='grey')

```

